---
title: "Statistical theory - FA2"
author: "Paul Joaquin M. Delos Santos"
date: "2024-09-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem Statement (3.49)

We are tasked with proving the equation:

$$
\sum_{j=1}^{N} (X_j - 1)^2 = \sum_{j=1}^{N} X_j^2 - 2 \sum_{j=1}^{N} X_j + N
$$

## Proof

Let us start by expanding the left-hand side of the equation:

$$
\sum_{j=1}^{N} (X_j - 1)^2
$$

### Step 1: Expanding the square

Expanding \( (X_j - 1)^2 \), we get:

$$
(X_j - 1)^2 = X_j^2 - 2X_j + 1
$$

Thus, we can rewrite the sum as:

$$
\sum_{j=1}^{N} (X_j - 1)^2 = \sum_{j=1}^{N} \left( X_j^2 - 2X_j + 1 \right)
$$

### Step 2: Splitting the sum

Now, break the summation into parts:

$$
\sum_{j=1}^{N} \left( X_j^2 - 2X_j + 1 \right) = \sum_{j=1}^{N} X_j^2 - 2 \sum_{j=1}^{N} X_j + \sum_{j=1}^{N} 1
$$

### Step 3: Simplifying the constant sum

Since the sum of \(1\) over \(N\) terms is just \(N\), we simplify this to:

$$
\sum_{j=1}^{N} 1 = N
$$

Thus, we now have:

$$
\sum_{j=1}^{N} (X_j - 1)^2 = \sum_{j=1}^{N} X_j^2 - 2 \sum_{j=1}^{N} X_j + N
$$

### Conclusion

Therefore, we have proven that:

$$
\sum_{j=1}^{N} (X_j - 1)^2 = \sum_{j=1}^{N} X_j^2 - 2 \sum_{j=1}^{N} X_j + N
$$

The equation holds, as required.



## Problem Statement (3.51)


```{r}
# Given values for U and V
U <- c(3, -2, 5)
V <- c(-4, -1, 6)

# (a) PUV = sum(U * V)
PUV <- sum(U * V)
PUV
# Explanation: This represents ∑U⋅V, the sum of the products of corresponding values of U and V.

# (b) P(U + 3)(V - 4) = sum((U + 3) * (V - 4))
PU_plus3_V_minus4 <- sum((U + 3) * (V - 4))
PU_plus3_V_minus4
# Explanation: This represents ∑(U + 3)(V - 4), the sum of the products after modifying U and V as described.

# (c) PV² = sum(V^2)
PV2 <- sum(V^2)
PV2
# Explanation: This is the sum of the squares of the values in V, ∑V².

# (d) (sum(U)) * (sum(V))^2
PU_sum <- sum(U)
PV_sum <- sum(V)
result_d <- PU_sum * (PV_sum^2)
result_d
# Explanation: Multiply the sum of U by the square of the sum of V, (∑U)(∑V)².

# (e) PUV² = sum(U * V^2)
PUV2 <- sum(U * V^2)
PUV2
# Explanation: This is ∑U⋅V², the sum of the products of U and the square of V.

# (f) sum(U^2 - 2*V^2 + 2)
result_f <- sum(U^2 - 2*V^2 + 2)
result_f
# Explanation: This calculates the sum after modifying U and V according to the given expression, ∑(U² - 2V² + 2).

# (g) sum(U / V)
result_g <- sum(U / V)
result_g
# Explanation: This computes the sum of the division of U by V, ∑(U/V).

```


## Problem Statement (3.51)
3.90 Find the geometric mean of the sets (a) 3, 5, 8, 3, 7, 2 and (b) 28.5, 73.6, 47.2, 31.5, 64.8.

$$
\text{Geometric Mean} = \left(\prod_{i=1}^{n} x_i\right)^{\frac{1}{n}}
$$

### Set (a)

The numbers are: 3, 5, 8, 3, 7, 2

```{r}
# Define the numbers for set (a)
set_a <- c(3, 5, 8, 3, 7, 2)

# Calculate the geometric mean for set (a)
geometric_mean_a <- prod(set_a)^(1/length(set_a))
geometric_mean_a

```
### Set (b)

The numbers are: 28.5, 73.6, 47.2, 31.5, 64.8

```{r}
# Define the numbers for set (b)
set_b <- c(28.5, 73.6, 47.2, 31.5, 64.8)

# Calculate the geometric mean for set (b)
geometric_mean_b <- prod(set_b)^(1/length(set_b))
geometric_mean_b

```


