---
title: "FA 5 - DELOS SANTOS"
author: "Paul Joaquin M. Delos Santos"
date: "2024-09-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Problem Overview
##  8.18 
Midwestern University has 1/3 of its students taking 9 credit hours, 1/3 taking 12 credit hours, and 1/3 taking 15 credit hours. 

Let \( X \) represent the credit hours a student is taking, with the distribution defined as:

- \( p(x) = \frac{1}{3} \) for \( x = 9, 12, 15 \)

### Mean and Variance of \( X \)

The mean \( \mu \) of \( X \) is calculated as:

\[
\mu = E(X) = \sum_{i} x_i p(x_i) = 9 \left(\frac{1}{3}\right) + 12 \left(\frac{1}{3}\right) + 15 \left(\frac{1}{3}\right) = 12
\]

The variance \( \sigma^2 \) of \( X \) is calculated as:

\[
\sigma^2 = E(X^2) - \mu^2
\]

Where \( E(X^2) \) is given by:

\[
E(X^2) = 9^2 \left(\frac{1}{3}\right) + 12^2 \left(\frac{1}{3}\right) + 15^2 \left(\frac{1}{3}\right) = 81 \left(\frac{1}{3}\right) + 144 \left(\frac{1}{3}\right) + 225 \left(\frac{1}{3}\right) = 150
\]

Thus, 

\[
\sigma^2 = 150 - 12^2 = 150 - 144 = 6
\]

The distribution of \( X \) is uniform across the three values.

### All Possible Samples of Size 2

We will now list all possible samples of size \( n = 2 \) from the population with replacement.


```{r}
# Define the population
population <- c(9, 12, 15)

# Generate all combinations of samples of size 2 with replacement
samples <- expand.grid(population, population)
samples
```

## Calculate Sample Means and Their Distribution



```{r}
# Calculate the sample means
sample_means <- rowMeans(samples)

# Calculate frequency of each mean
mean_freq <- as.data.frame(table(sample_means))
colnames(mean_freq) <- c("Sample_Mean", "Frequency")
mean_freq$Sample_Mean <- as.numeric(as.character(mean_freq$Sample_Mean))

# Calculate probabilities
mean_freq$p_xbar <- mean_freq$Frequency / sum(mean_freq$Frequency)

# Calculate xbar * p(xbar) and xbar^2 * p(xbar)
mean_freq$xbar_p <- mean_freq$Sample_Mean * mean_freq$p_xbar
mean_freq$xbar2_p <- (mean_freq$Sample_Mean^2) * mean_freq$p_xbar

# Show the results
mean_freq
```

## Plotting the Sampling Distribution of the Mean
```{r}
# Load necessary library
library(ggplot2)

# Plotting
ggplot(mean_freq, aes(x = Sample_Mean, y = Frequency)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Sampling Distribution of the Mean",
       x = "Sample Mean",
       y = "Frequency") +
  theme_minimal()

```

## 8.21
## Population Overview

The population consists of the following four numbers: \( 3, 7, 11, 15 \).

### (a) Calculate the Population Mean

The population mean \( \mu \) is calculated using the formula:

\[
\mu = \frac{\sum_{i=1}^{N} x_i}{N}
\]

Where \( N \) is the number of elements in the population. 

Given the numbers, the calculation will be:

\[
\mu = \frac{3 + 7 + 11 + 15}{4}
\]

### (b) Calculate the Population Standard Deviation

The population standard deviation \( \sigma \) is calculated using the formula:

\[
\sigma = \sqrt{\frac{\sum_{i=1}^{N} (x_i - \mu)^2}{N}}
\]

Substituting the calculated mean into the formula will yield the standard deviation.

### (c) Mean of the Sampling Distribution of Means

To find the mean of the sampling distribution of means for samples of size \( n = 2 \) drawn with replacement, we consider all possible combinations of samples.

Given the population \( \{3, 7, 11, 15\} \), the possible samples of size 2 are:

- \( (3, 3) \)
- \( (3, 7) \)
- \( (3, 11) \)
- \( (3, 15) \)
- \( (7, 3) \)
- \( (7, 7) \)
- \( (7, 11) \)
- \( (7, 15) \)
- \( (11, 3) \)
- \( (11, 7) \)
- \( (11, 11) \)
- \( (11, 15) \)
- \( (15, 3) \)
- \( (15, 7) \)
- \( (15, 11) \)
- \( (15, 15) \)

The sample means are then calculated for each combination. The mean of the sampling distribution of means is given by:

\[
\text{Mean of Sampling Distribution} = \frac{\sum \text{Sample Means}}{\text{Number of Samples}}
\]

### (d) Standard Deviation of the Sampling Distribution of Means

The standard deviation of the sampling distribution of means (standard error) can be calculated using the formula:

\[
\sigma_{\bar{x}} = \frac{\sigma}{\sqrt{n}}
\]

Where \( \sigma \) is the population standard deviation and \( n \) is the sample size. In this case, \( n = 2 \).


## 8.34
## Assuming there are 200 children born, we will calculate the following:

1. The probability that less than 40% will be boys.
2. The probability that between 43% and 57% will be girls.
3. The probability that more than 54% will be boys.

### Given Parameters

- Number of trials (children born), \( n = 200 \)
- Probability of having a boy or girl, \( p = 0.5 \)

### Mean and Standard Deviation

The mean (\( \mu \)) and standard deviation (\( \sigma \)) for the binomial distribution are calculated as follows:

- **Mean**:
\[
\mu = n \cdot p = 200 \cdot 0.5 = 100
\]

- **Standard Deviation**:
\[
\sigma = \sqrt{n \cdot p \cdot (1 - p)} = \sqrt{200 \cdot 0.5 \cdot 0.5} = \sqrt{50} \approx 7.07
\]

### (a) Probability that less than 40% will be boys

To find the probability that less than 40% of the 200 children will be boys:

- \( 40\% \) of 200 is \( 80 \) boys.

We calculate:
\[
P(X < 80)
\]

Using the normal approximation:
\[
Z = \frac{X - \mu}{\sigma} = \frac{80 - 100}{7.07} \approx -2.83
\]

Using the Z-table, we find:
\[
P(Z < -2.83) \approx 0.0023
\]

### (b) Probability that between 43% and 57% will be girls

This corresponds to between 43% and 57% boys:

- \( 43\% \) of 200 is \( 86 \) boys.
- \( 57\% \) of 200 is \( 114 \) boys.

We want:
\[
P(86 < X < 114)
\]

Calculating the Z-scores:
1. For \( 86 \):
   \[
   Z = \frac{86 - 100}{7.07} \approx -1.98
   \]

2. For \( 114 \):
   \[
   Z = \frac{114 - 100}{7.07} \approx 1.98
   \]

Using the Z-table:
\[
P(-1.98 < Z < 1.98) \approx P(Z < 1.98) - P(Z < -1.98) \approx 0.9761 - 0.0228 = 0.9533
\]

### (c) Probability that more than 54% will be boys

This corresponds to more than \( 54\% \) of 200, which is:

- \( 54\% \) of 200 is \( 108 \) boys.

We want:
\[
P(X > 108)
\]

Calculating the Z-score:
\[
Z = \frac{108 - 100}{7.07} \approx 1.13
\]

Using the Z-table:
\[
P(Z > 1.13) = 1 - P(Z < 1.13) \approx 1 - 0.8708 = 0.1292
\]

### Verification of Parts (c) and (d)

To verify the results for the mean and standard deviation of the sampling distribution:

1. **Mean of the Sampling Distribution**:
   - The mean of the sampling distribution of means should equal the population mean \( \mu \).
   - Therefore, we check if:
   \[
   \text{Mean of Sampling Distribution} = \mu
   \]

2. **Standard Deviation of the Sampling Distribution**:
   - The standard deviation of the sampling distribution \( \sigma_{\bar{x}} \) should equal the calculated standard error:
   \[
   \sigma_{\bar{x}} = \frac{\sigma}{\sqrt{n}}
   \]
   
## 8.49 
## Credit Hour Distribution at Metropolitan Technological College

The distribution of credit hours \( X \) and their corresponding probabilities \( p(x) \) are as follows:

| \( X \) | 6   | 9   | 12  | 15  | 18  |
|---------|-----|-----|-----|-----|-----|
| \( p(x) \) | 0.1 | 0.2 | 0.4 | 0.2 | 0.1 |

### Mean (\( \mu \)) and Variance (\( \sigma^2 \))

The mean \( \mu \) is calculated as:

\[
\mu = \sum_{i=1}^{N} x_i \cdot p(x_i)
\]

Calculating the mean:

\[
\mu = (6 \cdot 0.1) + (9 \cdot 0.2) + (12 \cdot 0.4) + (15 \cdot 0.2) + (18 \cdot 0.1)
\]

\[
\mu = 0.6 + 1.8 + 4.8 + 3.0 + 1.8 = 12
\]

The variance \( \sigma^2 \) is calculated as:

\[
\sigma^2 = \sum_{i=1}^{N} (x_i - \mu)^2 \cdot p(x_i)
\]

Calculating the variance:

1. Calculate each squared deviation from the mean and multiply by their respective probabilities:

\[
\sigma^2 = (6 - 12)^2 \cdot 0.1 + (9 - 12)^2 \cdot 0.2 + (12 - 12)^2 \cdot 0.4 + (15 - 12)^2 \cdot 0.2 + (18 - 12)^2 \cdot 0.1
\]

\[
= (36 \cdot 0.1) + (9 \cdot 0.2) + (0 \cdot 0.4) + (9 \cdot 0.2) + (36 \cdot 0.1)
\]

\[
= 3.6 + 1.8 + 0 + 1.8 + 3.6 = 10.8
\]

### Possible Samples of Size 2

We will consider all combinations of samples of size 2 drawn with replacement. The possible samples and their corresponding means are as follows:

| Sample   | Mean  | Probability |
|----------|-------|-------------|
| (6, 6)   | 6     | 0.1 * 0.1 = 0.01 |
| (6, 9)   | 7.5   | 0.1 * 0.2 = 0.02 |
| (6, 12)  | 9     | 0.1 * 0.4 = 0.04 |
| (6, 15)  | 10.5  | 0.1 * 0.2 = 0.02 |
| (6, 18)  | 12    | 0.1 * 0.1 = 0.01 |
| (9, 6)   | 7.5   | 0.2 * 0.1 = 0.02 |
| (9, 9)   | 9     | 0.2 * 0.2 = 0.04 |
| (9, 12)  | 10.5  | 0.2 * 0.4 = 0.08 |
| (9, 15)  | 12    | 0.2 * 0.2 = 0.04 |
| (9, 18)  | 13.5  | 0.2 * 0.1 = 0.02 |
| (12, 6)  | 9     | 0.4 * 0.1 = 0.04 |
| (12, 9)  | 10.5  | 0.4 * 0.2 = 0.08 |
| (12, 12) | 12    | 0.4 * 0.4 = 0.16 |
| (12, 15) | 13.5  | 0.4 * 0.2 = 0.08 |
| (12, 18) | 15    | 0.4 * 0.1 = 0.04 |
| (15, 6)  | 10.5  | 0.2 * 0.1 = 0.02 |
| (15, 9)  | 12    | 0.2 * 0.2 = 0.04 |
| (15, 12) | 13.5  | 0.2 * 0.4 = 0.08 |
| (15, 15) | 15    | 0.2 * 0.2 = 0.04 |
| (15, 18) | 16.5  | 0.2 * 0.1 = 0.02 |
| (18, 6)  | 12    | 0.1 * 0.1 = 0.01 |
| (18, 9)  | 13.5  | 0.1 * 0.2 = 0.02 |
| (18, 12) | 15    | 0.1 * 0.4 = 0.04 |
| (18, 15) | 16.5  | 0.1 * 0.2 = 0.02 |
| (18, 18) | 18    | 0.1 * 0.1 = 0.01 |

### Summary of Results

- Mean \( \mu = 12 \)
- Variance \( \sigma^2 = 10.8 \)


The table of samples, their means, and probabilities shows the distribution of the sample means for size 2 drawn with replacement from the population of credit hours.

