---
title: "FA7"
author: "Paul Joaquin M. Delos Santos"
date: "2024-09-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(dplyr)
library(car) # for Levene's Test
library(stats)


```



```{r }
# Create the dataset
data <- data.frame(
  Participant = 1:24,
  Cloak = c(rep(0, 12), rep(1, 12)),
  Mischief = c(3, 1, 5, 4, 6, 4, 6, 2, 0, 5, 4, 5,
               4, 3, 6, 6, 8, 5, 5, 4, 2, 5, 7, 5)
)
```

## Assumptions Checks
Assumption 1: Continuous Dependent Variable
The dependent variable (Mischief) is continuous.

# Assumption 2: Independent Groups
The independent variable (Cloak) consists of two categorical, independent groups.

# Assumption 3: Independence of Observations
Each participant belongs to only one group.

# Assumption 4: Outliers
```{r }
# Boxplot to inspect outliers
ggplot(data, aes(x = factor(Cloak), y = Mischief)) +
  geom_boxplot() +
  labs(title = "Boxplot of Mischief by Cloak Status",
       x = "Cloak Status (0 = Without, 1 = With)",
       y = "Number of Mischievous Acts")
```

# Assumption 5: Normality
```{r}
# Shapiro-Wilk test for normality
shapiro_test_without_cloak <- shapiro.test(data$Mischief[data$Cloak == 0])
shapiro_test_with_cloak <- shapiro.test(data$Mischief[data$Cloak == 1])

shapiro_test_without_cloak
shapiro_test_with_cloak


```

# Assumption 6: Homogeneity of Variances
```{r}
# Levene's Test for homogeneity of variances
levene_test_result <- leveneTest(Mischief ~ factor(Cloak), data = data)
levene_test_result




```


# Comparison of two means 
```{r}

# Comparison of Two Means 
# Calculate group means
library(dplyr) # Ensure dplyr is loaded
group_means <- data %>%
  group_by(Cloak) %>%
  summarise(Mean_Mischief = mean(Mischief), .groups = 'drop')

# Print group means
print(group_means)

# Simple bar plot to visualize means
ggplot(group_means, aes(x = factor(Cloak), y = Mean_Mischief, fill = factor(Cloak))) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = round(Mean_Mischief, 2)), vjust = -0.5) +
  labs(title = "Comparison of Mischief by Cloak Status",
       x = "Cloak Status (0 = Without, 1 = With)",
       y = "Mean Number of Mischievous Acts") +
  theme_minimal()



```



 rain cloud plot "# Raincloud plot to visualize means
 
```{r}
ggplot(data, aes(x = factor(Cloak), y = Mischief, fill = factor(Cloak))) +
  geom_violin(trim = TRUE, alpha = 0.5) +
  geom_jitter(aes(color = factor(Cloak)), width = 0.1, alpha = 0.5) +
  labs(title = "Raincloud Plot of Mischief by Cloak Status",
       x = "Cloak Status (0 = Without, 1 = With)",
       y = "Number of Mischievous Acts") +
  theme_minimal() +
  theme(legend.position = "none")


  
  
```
  
  
# Independent Samples T-Test
```{r}
# Conducting the independent samples t-test
t_test_result <- t.test(Mischief ~ Cloak, data = data)
t_test_result



```
