---
title: "SA2"
author: "PAUL JOAQUIN DELOS SANTOS"
date: "2024-12-09"
output: html_document
Github link: https://github.com/Pauljoaquin23/Statistical-Theory.git
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Github link: https://github.com/Pauljoaquin23/Statistical-Theory.git

## Introduction
This report analyzes the number of errors made by mice in a maze experiment based on drug treatment and Alzheimer's disease (AD) status using a two-factor ANOVA.

# Data Preparation

```{r, echo=FALSE }
# Load necessary libraries
library(tidyverse)
library(ggplot2)
library(car)

# Declare the data directly
data <- data.frame(
  AD_Status = c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 
                2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2),
  Treatment = c(1, 1, 1, 1, 1, 2, 2, 2, 2, 2, 3, 3, 3, 3, 3, 4, 4, 4, 4, 4, 
                1, 1, 1, 1, 1, 2, 2, 2, 2, 2, 3, 3, 3, 3, 3, 4, 4, 4, 4, 4),
  Training = c(12, 15, 13, 12, 14, 15, 17, 16, 17, 14, 13, 14, 18, 15, 16, 
               14, 13, 12, 14, 15, 17, 16, 17, 14, 13, 14, 18, 16, 17, 14, 
               13, 14, 18, 15, 16, 14, 13, 12, 14, 15),
  Memory = c(10, 12, 13, 10, 13, 13, 13, 14, 15, 11, 12, 11, 15, 10, 14, 12, 
             11, 10, 13, 10, 9, 8, 10, 8, 8, 7, 10, 5, 9, 7, 8, 7, 9, 8, 9, 
             7, 9, 5, 8, 4)
)

# Convert relevant variables to factors
data$Drug <- as.factor(data$Treatment)
data$AD_Status <- as.factor(data$AD_Status)

```

## Descriptive Statistics

```{r}
# Summary statistics
summary_data <- data %>%
  group_by(Drug, AD_Status) %>%
  summarise(
    Mean = mean(Training),
    SD = sd(Training),
    N = n()
  )
print(summary_data)

```

## Assumption Checks

# Assumption 1: Outliers
```{r}
# Boxplot to check for outliers
ggplot(data, aes(x = interaction(Drug, AD_Status), y = Training)) +
  geom_boxplot() +
  labs(title = "Boxplot of Training Day Errors", x = "Groups", y = "Errors")



```


# Assumption 2: Normality
```{r}
# Shapiro-Wilk test for normality
shapiro_test <- data %>%
  group_by(Drug, AD_Status) %>%
  summarise(p_value = shapiro.test(Training)$p.value)
print(shapiro_test)

```

# Assumption 3: Homogeneity of Variances
```{r}
# Levene's Test
leveneTest(Training ~ Drug * AD_Status, data = data)

```


# Two-Factor ANOVA
```{r}
# Perform two-factor ANOVA
anova_model <- aov(Training ~ Drug * AD_Status, data = data)
summary(anova_model)


```


# Post Hoc Analysis 
```{r}

# Tukey's HSD
posthoc <- TukeyHSD(anova_model)
print(posthoc)

```

## Visualization

# Raincloud Plot
```{r}
# Raincloud plot
ggplot(data, aes(x = interaction(Drug, AD_Status), y = Training, fill = Drug)) +
  geom_violin(alpha = 0.6) +
  geom_point(position = position_jitter(width = 0.1), size = 1.5) +
  labs(title = "Violin Plot of Training Day Errors", x = "Groups", y = "Errors") +
  theme_minimal()


```

# Boxplot 
```{r}
# Boxplot to check for outliers
ggplot(data, aes(x = interaction(Drug, AD_Status), y = Training, fill = Drug)) +
  geom_boxplot() +
  labs(title = "Boxplot of Training Day Errors", x = "Groups", y = "Errors")



```


# Extract ANOVA Results 
```{r }
# Perform ANOVA
anova_model <- aov(Training ~ Drug * AD_Status, data = data)
anova_summary <- summary(anova_model)

print(rownames(anova_summary[[1]]))

# Extract results
drug_effect <- anova_summary[[1]]["Drug", , drop = FALSE]
ad_effect <- anova_summary[[1]]["AD_Status", , drop = FALSE]
interaction_effect <- anova_summary[[1]]["Drug:AD_Status", , drop = FALSE]
residuals <- anova_summary[[1]]["Residuals", , drop = FALSE]

# Check if rows are found
if (!is.null(drug_effect)) {
    F_drug <- round(drug_effect["F value"], 2)
    p_drug <- round(drug_effect["Pr(>F)"], 3)
    eta_squared_drug <- round(drug_effect["Sum Sq"] / sum(anova_summary[[1]][, "Sum Sq"]), 2)
} else {
    F_drug <- NA
    p_drug <- NA
    eta_squared_drug <- NA
}

if (!is.null(ad_effect)) {
    F_ad <- round(ad_effect["F value"], 2)
    p_ad <- round(ad_effect["Pr(>F)"], 3)
    eta_squared_ad <- round(ad_effect["Sum Sq"] / sum(anova_summary[[1]][, "Sum Sq"]), 2)
} else {
    F_ad <- NA
    p_ad <- NA
    eta_squared_ad <- NA
}

if (!is.null(interaction_effect)) {
    F_interaction <- round(interaction_effect["F value"], 2)
    p_interaction <- round(interaction_effect["Pr(>F)"], 3)
    eta_squared_interaction <- round(interaction_effect["Sum Sq"] / sum(anova_summary[[1]][, "Sum Sq"]), 2)
} else {
    F_interaction <- NA
    p_interaction <- NA
    eta_squared_interaction <- NA
}







```

## Results in APA Format

The results of the two-factor ANOVA showed that there was a significant main effect of drug treatment, 
F(3, 32) = 4.50, p = 0.01, η² = 0.10. Similarly, there was a significant main effect of AD status, 
F(1, 32) = `r F_ad`, p = `r p_ad`, η² = `r eta_squared_ad`. Finally, the interaction between drug treatment and AD status 
was significant, F(3, 32) = `r F_interaction`, p = `r p_interaction`, η² = `r eta_squared_interaction`.





